Software Product Line Analysis Techniques

### 2018-01-09 doc/quals/quals.tex
Abstract Interpretation

\begin{abstract}
TODO
\end{abstract}

@@

This lifting approach scales well with the number of features. Complexity grows polynomially with the number of statements in an SPL, rather than exponentially in the number of features. In addition, the implementation takes the feature model of the SPL into consideration, pruning statements with invalid presence conditions. This eliminates the need to do excessive computations that are of no value to the SPL analysis.

This work was evaluated on 3 lifted analyses: Possible Types, Reaching Definitions and Unititialized Variables. 4 real world SPLs were used as case studies: BerkeleyDB, GPL, Lampiro and MM08. Performance results show that the lifted analyses run in minutes (in some cases seconds), while brute force analysis would take years for some of the case studies. One thing to note though is that the correctness of the lifting approach is shown empirically rather than with a logical proof.

@@

\section{Variability-Aware Abstract Interpretation}
\cite{Midtgaard:2015}

@@

Their family-based approach is more interesting. Features are encoded as boolean variables, and feature expressions as Binary Decision Digarams (BDDs). Variability within a source-code function is mapped into multiple implementations, with an additional dispatcher function choosing which implementation to call based on feature expressions. This simply reduces variability into conditionals, allowing existing model checkers to be used as is.

@@

Similar to \cite{Apel:2013}, the main principles guiding the design of the variability-aware analyses are late splitting and early joining. These principles maximize reuse of intermediate analysis information across different variants, eventually resulting in variability-aware analyses outperformaing brute-force per-product analysis by orders of magnitude. A third principle implicit in \cite{Apel:2013} but explicitly spelled out here is local variability representation. This is the mechanism of maintaining variability within intermediate analysis values.

@@

\section{SuperC}
In \cite{Gazzillo:2012}, a variability-aware C-language parser and preprocessor are presented and evaluated against a similar C parser implemented as a part of the TypeChef project \cite{Kastner:2011}. SuperC includes a lexer, a preprocessor and a Fork Merge LR (FMLR) parser. Several optimizations have been implemented to lower the number of forked subparsers, and consequently speedup parsing. On the x86 Linux kernel codebase, SuperC is 3.4 to 3.8 times faster than TypeChef.

The idea behind FMLR parsing is to fork subparsers (initially sharing the state of their parent) on each variability point. A naive implementation would result in a number of subparsers exponential in the number of variability features. To prevent this exponential explosion, SuperC implements 4 optimizations: follow-set calculation, early reduces, lazy shifts, and shared reduces. With those optimizations, for almost all the Linux kernel compilation units the number of subparsers is less than 20.

>>>>

%\begin{abstract}
%TODO
%\end{abstract}

@@

This lifting approach scales well with the number of features. Complexity grows polynomially with the number of statements in an SPL, rather than exponentially in the number of features. In addition, the implementation takes the feature model of the SPL into consideration, pruning statements with invalid presence conditions. This eliminates the need to do excessive computations that are of no value to the SPL analysis. Presence conditions are encoded as BDDs, which further optimizes their satisfiability checks.

This work was evaluated on 3 lifted analyses: Possible Types, Reaching Definitions and Uninitialized Variables. 4 real world SPLs were used as case studies: BerkeleyDB, GPL, Lampiro and MM08. Performance results show that the lifted analyses run in minutes (in some cases seconds), while brute force analysis would take years for some of the case studies. One thing to note though is that the correctness of the lifting approach is shown empirically rather than with a logical proof.

@@

\section{Variability-Aware Abstract Interpretation}

In \cite{Midtgaard:2015}, a systematic approach to lift abstract interpretation analyses to SPLs is presented. The paper provides a rigorous  formal treatment of abstract interpretation applied to a simple imperative language, with variability at the statement level. Formal proofs for the correctness of the lifting approach are also presented.

The lifting approach provides systematic liftings of the abstract domains, Galois connections, and corresponding analyses by the use of combinators. It is worth noting that this approach is systematic but not automated. Language-specific theorems need to be proven (3 of them have been proven in the paper), in addition to a language-agnostic framework theorem that has been proven as well.

The paper also presents some optimizations that would trade analysis precision for speed. The idea is to allow for variational abstraction, which would further abstract variability. Further optimization techniques involve sharing analysis intermediate values while keeping track of presence conditions (either in CNF/DNF formulas or BDDs). In addition, a technique for extracting data-flow equations (both the product-level equations and the lifted equations) is explained.

This work was evaluated on a Reaching Definitions analysis for Java programs. Three Java SPLs were used: GPL, Prevayler and BerkeleyDB. The speedup due to the different optimizations for the lifting approach (with respect to brute force) is exponential with respect to the number of SPL configurations.

@@

Their family-based approach is more interesting. Features are encoded as boolean variables, and feature expressions as Binary Decision Diagrams (BDDs). Variability within a source-code function is mapped into multiple implementations, with an additional dispatcher function choosing which implementation to call based on feature expressions. This simply reduces variability into conditionals, allowing existing model checkers to be used as is.

@@

Similar to \cite{Apel:2013}, the main principles guiding the design of the variability-aware analyses are late splitting and early joining. These principles maximize reuse of intermediate analysis information across different variants, eventually resulting in variability-aware analyses outperforming brute-force per-product analysis by orders of magnitude. A third principle implicit in \cite{Apel:2013} but explicitly spelled out here is local variability representation. This is the mechanism of maintaining variability within intermediate analysis values.

@@

\section{SuperC}
In \cite{Gazzillo:2012}, a variability-aware C-language parser and preprocessor are presented and evaluated against a similar C parser implemented as a part of the TypeChef project \cite{Kastner:2011}. SuperC includes a lexer, a preprocessor and a Fork Merge LR (FMLR) parser. Several optimizations have been implemented to lower the number of forked sub-parsers, and consequently speedup parsing. On the x86 Linux kernel codebase, SuperC is 3.4 to 3.8 times faster than TypeChef.

The idea behind FMLR parsing is to fork sub-parsers (initially sharing the state of their parent) on each variability point. A naive implementation would result in a number of sub-parsers exponential in the number of variability features. To prevent this exponential explosion, SuperC implements 4 optimizations: follow-set calculation, early reduces, lazy shifts, and shared reduces. With those optimizations, for almost all the Linux kernel compilation units the number of sub-parsers is less than 20.
